Ae albopictus CPP manuscript

### 2019-10-25 MS.Rmd
MS edits

---
title: "Ae albopictus CPP manuscript"
output:
  word_document: 
    fig_width: 3.5
    fig_height: 3
  html_document: default
---

@@

```{r DI_plot, echo=FALSE, include=FALSE}
#plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)

@@

```{r Japan_US_years, echo=FALSE, include=FALSE}
ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
  scale_color_manual(values=c("white","gray","black")) + 
  scale_fill_manual(values=c("white","gray","black"))+ 
  geom_point(shape=21, color="black") + 
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  facet_grid(~country) + stat_smooth(method=lm, se=FALSE)
```

@@

```{r, echo=FALSE, include=FALSE}
ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) 
```

@@

country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none")
```

@@

DI_pies_2018<-plot_grid(remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), ncol=1, align="hv")
```

@@

- Diapause is an important overwintering adaptation in many organisms
- This adaptation comes with trade-offs, but exactly what they are is unclear (winter phenotypes of diapausing organisms; life history strategies)
- Climate change may have unpredictable effects on growing seasons
- Invasive species provide natural experiments for rapid adaptation
- Aedes albopictus worldwide spread: diapause played a role 
- What is the continuing role of diapause in light of environmental change, continuing introductions/admixture,

@@

### Mosquito population sampling and rearing
Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from (Fig 1). Sampling locations are described in Table 1. Locations in the US were largely identical to locations sampled in 2008. (More details about field sampling.) At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had X females and Y males.

@@

### Statistical analysis
Critical photoperiod was calculated from photoperiod response curves as ED50 using a five-parameter generalized log-logistic model in the R (3.6.0) package drc (3.0.1). Effects of latitude, country and year (2008 vs. 2018) on CPP and on DI were analyzed with linear models. Additional packages used for calculations, analysis and figures included tidyverse (1.2.1), maps, mapdata, scatterpie…(update package list as figures come together).

@@

Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 1, Table 2). Diapause incidence was high (generally greater than 90%) at latitudes above ~32.5 degrees in both countries (Figure 2A). Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan. Diapause incidence followed a latitudinal cline in the US, but a cline could not be observed in Japan due to absence of land to sample between Okinawa and Kagoshima. In the limited number of southern populations that could be compared between the US and Japan, latitude-specific diapause incidence appeared to be lower in the US (Figure 2B).

```{r fig1, fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE}
incidence_lat_year_plot
```

@@

As observed in

>>>>

---
title: "Ae albopictus CPP manuscript"
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
output:
  pdf_document: default
  html_document: default
  word_document: default
---

@@

```{r DI_plot, echo=FALSE, include=FALSE}
#plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)

@@

```{r Japan_US_years, echo=FALSE, include=FALSE}
cpp_japan_us_years<-ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
  scale_color_manual(values=c("white","gray","black")) + 
  scale_fill_manual(values=c("white","gray","black"))+ 
  geom_point(shape=21, color="black") + 
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  facet_grid(~country) + stat_smooth(method=lm, se=FALSE)
```

@@

```{r, echo=FALSE, include=FALSE}
cpp_08_18_shifts<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
  geom_point() + scale_color_manual(values=c("red", "blue")) +
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + theme_bw() + theme(legend.position=c(.8,.25))
```

@@

country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm"))
```

@@

DI_pies_2018<-plot_grid(remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), ncol=1, align="hv")
```

@@

- Diapause is an important overwintering adaptation in many organisms
- This adaptation comes with trade-offs, but exactly what they are is unclear (winter phenotypes of diapausing organisms; life history strategies)
- Climate change may have unpredictable effects on growing seasons
- Invasive species provide natural experiments for rapid adaptation
- Aedes albopictus worldwide spread: diapause played a role 
- What is the continuing role of diapause in light of environmental change, continuing introductions, competing selective pressures? 
- We recapitulated 2008 CPP study

@@

### Mosquito population sampling and rearing
Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`. Sampling locations are described in Table 1. Locations in the US were largely identical to locations sampled in 2008. [More details about field sampling.] At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had X females and Y males.

@@

Photoperiod response data from 2008 and calculated CPP values from 1988 were obtained from Urbanski et al 2012 data files (doi:10.5061/dryad.68277435).

### Statistical analysis
Critical photoperiod was calculated from photoperiod response curves (Figure S1) as ED50 using a five-parameter generalized log-logistic model in the R (3.6.0) package drc (`r packageVersion("drc")`). Effects of latitude, country and year (2008 vs. 2018) on CPP and on DI were analyzed with linear models. Additional packages used for calculations, analysis and figures included tidyverse (`r packageVersion("tidyverse")`), maps (`r packageVersion("maps")`), mapdata (`r packageVersion("mapdata")`), scatterpie (`r packageVersion("scatterpie")`)…[update package list as figures come together].

@@

Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 1, Table 2).  Diapause incidence was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries (Figure 2A). Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan. Diapause incidence followed a latitudinal cline in the US, but a cline could not be observed in Japan due to absence of land to sample between Okinawa and Kagoshima. In the limited number of southern populations that could be compared between the US and Japan, latitude-specific diapause incidence appeared to be lower in the US (Figure 2B).

```{r fig1, fig.pos="H", fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE}
incidence_lat_year_plot
```

@@

CPP values correlated strongly with latitude in the US and Japan. Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3).

```{r fig3, fig.cap="Figure 3: Critical photoperiod (CPP) as a function of latitude in the US and Japan in 1988, 2008, and 2018. Error bars represent standard errors of CPP estimates. Errors around 1988 CPP estimates are absent because experiment and calculations were conducted using different methods and raw data were not available.", echo=FALSE, message=FALSE}
cpp_japan_us_years
```

```{r table3, echo=FALSE}
kable(anova(cppmodel), caption="Table 3: Effects of latitude, country and year on critical photoperiod in 2008 and 2018.")
```

However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (Figure 4). Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008. This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.

```{r fig4, fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions; dotted=2018, solid=2008.", echo=FALSE}
cpp_08_18_shifts
```

## Supplementary Figures
```{r figS1, fig.cap="Figure S1: Photoperiod response curves", echo=FALSE}
DI_plot
```
